## Data Analysis - Results
### Summary
We found a significant decrease in 5-hydroxi-methyl-cytosine (5hmC) expression from normal to tumor tissues (P < 0.001), although differences were not apparent between noninvasive and invasive tumors (P = 0.60). We did not found significant associations between 5hmC expression and clinicopathologic features and outcome. Kaplan-Meier curves were not significantly different for patients with low or high 5hmC expression regarding overall mortality and cancer-related mortality. Logistic regression models and Cox's proportional hazard models did not show evidence of a prognostic usefulness of 5hmC in predicting outcome, either in unadjusted or in adjusted models.

***

### Global Options and Dataset
* Defining global options
```{r global}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, fig.height = 7, fig.width = 9)
```

* Loading the dataset, libraries and sources for analysis
```{r load}
data <- read.delim("DatasetBladder.csv")
attach(data)
library(survival)
source("../StandardAnalysis2.R")
source("../SurvivalPlot.R")
```

***

### Description of the Dataset
In this section we will describe the dataset, including patient's clinical features, preoperative and surgical data, pathologic features of the tumor, as well as preoperative and posoperative treatment, and patient's outcome. For this section we will use the full dataset (N = `r nrow(data)`). This dataset include all cases that were used for building the tissue microarrays. Data for this section were obtain from medical charts and pathology reports.

#### Patient's clinical features
```{r clinical}
# Patient's age
descriptive(age)
# Patient's sex
table.prop(sex)
# Patient's race
table.prop(race)
# Concomitant prostate carcinoma
table.prop(pca)
```
#### Preoperative data
```{r preoperative}
# Preoperative specimen
table.prop(pre.specimen)
# Preoperative pT stage
table.prop(pre.pt)
# Preoperative clinical stage
table.prop(pre.stage)
# Presence of CIS in preoperative specimen
table.prop(pre.cis)
```
#### Surgical data
```{r surgical}
# Surgical specimen
table.prop(specimen)
# pT stage at cystectomy
table.prop(pt.cystectomy)
# Clinical stage at cystectomy
table.prop(stage.cystectomy)
```
#### Pathologic data
```{r}
# Diagnosis according to the pathology report
table.prop(histo.dx)
# Diagnosis at the TMA spot
table.prop(spot.dx)
# Histologic type of the tumor
table.prop(type)
# Presence of CIS
table.prop(cis)
# Lymphovascular invasion
table.prop(lvi)
# Total number of lymph nodes
descriptive(nodes.total)
# Presence of lymph nodes metastasis
table.prop(ln.met)
# Total number of metastatic lymph nodes
descriptive(nodes.positive)
```
#### Treatment
```{r treatment}
# Preoperative radiotherapy
table.prop(pre.rt)
# Preoperative chemotherapy
table.prop(pre.ct)
# Intravesical chemotherapy
table.prop(intra.ct)
# Intravesical chemotherapy type
table.prop(intra.ct.type)
# Posoperative radiotherapy
table.prop(pos.rt)
# Posoperative chemotherapy
table.prop(pos.ct)
# Posoperative chemotherapy type
table.prop(pos.ct.type)
```
#### Outcome
```{r}
# Follow-up from cystectomy, in months
descriptive(fu.cystectomy)
# Follow-up from diagnosis, in months
descriptive(fu.diagnosis)
# Early complications
table.prop(comp.early)
# Late complications
table.prop(comp.late)
# Metastatic disease after cystectomy
table.prop(metastasis)
# Tumor progression, i.e., metastatic disease or patient dying of cancer
table.prop(progression)
# Overall mortality, i.e., patient dying of any cause
table.prop(death)
# Cancer mortality, i.e., patient dying of cancer (No/Yes)
table.prop(dod)
```
#### Overall 5-hydroxi-methyl-cytosine expression
```{r 5hmc}
# Tumor 5hmC median H-score
descriptive(t.5hmc.med)
# Normal 5hmC median H-score
descriptive(n.5hmc.med)
# Tumor 5hmC mean H-score
descriptive(t.5hmc.m)
# Normal 5hmC mean H-score
descriptive(n.5hmc.m)
```

***

### 5hmC expression in nontumor and tumor tissue
In this section we will compare 5hmC H-scores as they were estimated in paired (i.e, form the sama surgical specimen) normal (i.e., histologically-normal urothelium) and tumor (both invasive and noninvasive) tissues. For this we will use the full dataset. Comparisons will be made using the Mann-Whitney U test.

* Mean H-scores
```{r meanComparisonNormal}
# Comparing mean H-scores in nontumor and tumor tissues
compare.mw2(n.5hmc.m, t.5hmc.m)
# Ploting a boxplot
boxplot(n.5hmc.m, t.5hmc.m, notch = TRUE, 
        ylab = "Mean 5hmC H-scores", cex.lab = 1.2,
        col = gray.colors(2),
        names = c("Normal Cells", "Tumor Cells")
        )
# Plotting a violin plot
require(vioplot)
vioplot(n.5hmc.m, t.5hmc.m,
        col = "gray",
        rectCol = "white", colMed = "black",
        names = c("Normal Cells", "Tumor Cells")
        )
title(ylab = "Mean 5hmC H-scores", cex.lab = 1.2)
```

* Median H-scores
```{r medianComparisonNormal}
# Comparing median H-scores in nontumor and tumor tissues
compare.mw2(n.5hmc.med, t.5hmc.med)
# Ploting a boxplot
boxplot(n.5hmc.med, t.5hmc.med, notch = TRUE, 
        ylab = "Median 5hmC H-scores", cex.lab = 1.2,
        col = gray.colors(2),
        names = c("Normal Cells", "Tumor Cells")
        )
# Plotting a violin plot
require(vioplot)
vioplot(n.5hmc.med, t.5hmc.med,
        col = "gray",
        rectCol = "white", colMed = "black",
        names = c("Normal Cells", "Tumor Cells")
        )
title(ylab = "Median 5hmC H-scores", cex.lab = 1.2)
```

In addition, we will compare mean and median 5hmC scores between noninvasive and invasive tumors. Mean and median H-scores will be obtained from `t.5hmc.m` and `t.5hmc.med` (respectively), using `spot.dx` to identify which cases corresponded to noninvasive or invasive tumors. Regarding the latter, in some cases there was a discrepancy between the diagnosis given in the pathology report (as specified in `histo.dx`) and what was found at the TMA spot (specified in `spot.dx`). The following table shows what was found at the TMA spots.
```{r spotDx}
table.prop(spot.dx)
```
Thus, for the next comparisons and plots we will use only the `Invasive` and `Noninvasive` cases.
* Mean H-scores
```{r meanComparisonTumor}
# Selecting only invasive and noninvasive cases from spot.dx
mean.5hmc.invasive <- t.5hmc.m[spot.dx == "Invasive"]
mean.5hmc.noninvasive <- t.5hmc.m[spot.dx == "Noninvasive"]
# Comparing median H-scores in nontumor and tumor tissues
compare.mw2(mean.5hmc.invasive, mean.5hmc.noninvasive)
# Ploting a boxplot
boxplot(mean.5hmc.invasive, mean.5hmc.noninvasive, notch = TRUE, 
        ylab = "Mean 5hmC H-scores", cex.lab = 1.2,
        col = gray.colors(2),
        names = c("Invasive Tumors", "Noninvasive Tumors")
        )
# Plotting a violin plot
require(vioplot)
vioplot(mean.5hmc.invasive, mean.5hmc.noninvasive,
        col = "gray",
        rectCol = "white", colMed = "black",
        names = c("Invasive Tumors", "Noninvasive Tumors")
        )
title(ylab = "Mean 5hmC H-scores", cex.lab = 1.2)
```

* Median H-scores
```{r medianComparisonTumor}
# Selecting only invasive and noninvasive cases from spot.dx
median.5hmc.invasive <- t.5hmc.med[spot.dx == "Invasive"]
median.5hmc.noninvasive <- t.5hmc.med[spot.dx == "Noninvasive"]
# Comparing median H-scores in nontumor and tumor tissues
compare.mw2(median.5hmc.invasive, median.5hmc.noninvasive)
# Ploting a boxplot
boxplot(median.5hmc.invasive, median.5hmc.noninvasive, notch = TRUE, 
        ylab = "Median 5hmC H-scores", cex.lab = 1.2,
        col = gray.colors(2),
        names = c("Invasive Tumors", "Noninvasive Tumors")
        )
# Plotting a violin plot
require(vioplot)
vioplot(median.5hmc.invasive, median.5hmc.noninvasive,
        col = "gray",
        rectCol = "white", colMed = "black",
        names = c("Invasive Tumors", "Noninvasive Tumors")
        )
title(ylab = "Median 5hmC H-scores", cex.lab = 1.2)
```

***

### 5hmC expression and outcome
For this section we will use 3 different approaches for evaluating the association of 5hmC and outcome: 1) using 5hmC H-scores; 2) categorizing 5hmC expression in low/high levels using the median as the cutoff point; 3) categorizing 5hmC expression in low/high levels using the Q3 as the cutoff point.

Only invasive cases will be consider from now on, as observed at the TMA spot (i.e., `spot.dx == Invasive`).
```{r invasive}
invasive <- subset(data, spot.dx == "Invasive")
detach(data)
attach(invasive)
```
The number of invasive cases is N = `r nrow(invasive)`. Also, only median H-scores will be consider for analysis.

#### Approach #1
In this approach we will compare 5hmC `H-scores` against clinicopathologic features and outcome. Comparisons will be made using the Mann-Whitney U test. For age (numeric variable) we will use the median to define younger and older age groups.
```{r outcome1}
# 5hmC and patient's age
age.group <- ifelse(age > median(age), c("Older age"), c("Younger age"))
compare.mw(t.5hmc.med, age.group)
# 5hmC and patient's sex
compare.mw(t.5hmc.med, sex)
# 5hmC and patient's race
compare.mw(t.5hmc.med, race)
# 5hmC and pT stage at cystectomy
compare.kw(t.5hmc.med, pt.cystectomy)
# 5hmC and clinical stage at cystectomy
compare.kw(t.5hmc.med, stage.cystectomy)
# 5hmC and lymphovascular invasion
compare.mw(t.5hmc.med, lvi)
# 5hmC and lymph node metastasis
compare.mw(t.5hmc.med, ln.met)
# 5hmC and early complications
compare.mw(t.5hmc.med, comp.early)
# 5hmC and late complications
compare.mw(t.5hmc.med, comp.late)
# 5hmC and metastatic disease after cystectomy
compare.mw(t.5hmc.med, metastasis)
# 5hmC and overall mortality
compare.mw(t.5hmc.med, death)
# 5hmC and cancer-related mortality
compare.mw(t.5hmc.med, dod)
# 5hmC and tumor progression
compare.mw(t.5hmc.med, progression)
```

#### Approach # 2
In this approach we will compare 5hmC levels (categorizing 5hmC expression in low/high levels using the `median` as the cutoff point) against clinicopathologic features and outcome. Comparisons will be made using the Fisher's exact test. For age (numeric variable) we will use the median to define younger and older age groups.
```{r outcome2}
# Categorizing 5hmC expression
med.5hmc <- ifelse(t.5hmc.med > median(t.5hmc.med), c("High"), c("Low"))
# 5hmC and patient's age
compare.fisher(med.5hmc, age.group)
# 5hmC and patient's sex
compare.fisher(med.5hmc, sex)
# 5hmC and patient's race
compare.fisher(med.5hmc, race)
# 5hmC and pT stage at cystectomy
compare.fisher(med.5hmc, pt.cystectomy)
# 5hmC and clinical stage at cystectomy
compare.fisher(med.5hmc, stage.cystectomy)
# 5hmC and lymphovascular invasion
compare.fisher(med.5hmc, lvi)
# 5hmC and lymph node metastasis
compare.fisher(med.5hmc, ln.met)
# 5hmC and early complications
compare.fisher(med.5hmc, comp.early)
# 5hmC and late complications
compare.fisher(med.5hmc, comp.late)
# 5hmC and metastatic disease after cystectomy
compare.fisher(med.5hmc, metastasis)
# 5hmC and overall mortality
compare.fisher(med.5hmc, death)
# 5hmC and cancer-related mortality
compare.fisher(med.5hmc, dod)
# 5hmC and tumor progression
compare.fisher(med.5hmc, progression)
```

#### Approach # 3
In this approach we will compare 5hmC levels (categorizing 5hmC expression in low/high levels using the `upper tertile` as the cutoff point) against clinicopathologic features and outcome. Comparisons will be made using the Fisher's exact test. For age (numeric variable) we will use the median to define younger and older age groups.
```{r outcome3}
# Categorizing 5hmC expression
q3.5hmc <- ifelse(t.5hmc.med > quantile(t.5hmc.med, .75, na.rm = TRUE), c("High"), c("Low"))
# 5hmC and patient's age
compare.fisher(q3.5hmc, age.group)
# 5hmC and patient's sex
compare.fisher(q3.5hmc, sex)
# 5hmC and patient's race
compare.fisher(q3.5hmc, race)
# 5hmC and pT stage at cystectomy
compare.fisher(q3.5hmc, pt.cystectomy)
# 5hmC and clinical stage at cystectomy
compare.fisher(q3.5hmc, stage.cystectomy)
# 5hmC and lymphovascular invasion
compare.fisher(q3.5hmc, lvi)
# 5hmC and lymph node metastasis
compare.fisher(q3.5hmc, ln.met)
# 5hmC and early complications
compare.fisher(q3.5hmc, comp.early)
# 5hmC and late complications
compare.fisher(q3.5hmc, comp.late)
# 5hmC and metastatic disease after cystectomy
compare.fisher(q3.5hmc, metastasis)
# 5hmC and overall mortality
compare.fisher(q3.5hmc, death)
# 5hmC and cancer-related mortality
compare.fisher(q3.5hmc, dod)
# 5hmC and tumor progression
compare.fisher(q3.5hmc, progression)
```

***

### Survival plots
In this section we will plot survival curves for overall mortality and cancer-related mortality according to 5hmC expression, categorizing 5hmC expression in low/high levels using as cutoff points: 1) the median; 2) the upper tertile. We will use `follow-up from cystectomy` for the time-to-event analysis.

* Defining survival objects
```{r survobj}
death.surv <- Surv(fu.cystectomy, as.numeric(death))
dod.surv <- Surv(fu.cystectomy, as.numeric(dod))
```

#### Approach #1
In this approch we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality using the `median` H-score as the cutoff point for defining low/high 5hmC levels. We will compare survival curves using the `log-rank test`.

* Overall mortality
```{r kmplots1A}
# Estimating P values using the log-rank test
death.med.lr <- survdiff(death.surv ~ med.5hmc)
death.med.p <- pchisq(death.med.lr$chisq, df = 1, lower = FALSE)
# Plotting curves
survival.plot(x = survfit(death.surv ~ med.5hmc), title = "Overall Mortality and 5hmC Expression", levels = c("High", "Low"))
text(paste("P (log-rank test) = ", round(death.med.p, digits = 3)), x = 35, y = 0.01, cex = 1.25)
```

* Cancer-related mortality
```{r kmplots1B}
# Estimating P values using the log-rank test
dod.med.lr <- survdiff(dod.surv ~ med.5hmc)
dod.med.p <- pchisq(dod.med.lr$chisq, df = 1, lower = FALSE)
# Plotting curves
survival.plot(x = survfit(dod.surv ~ med.5hmc), title = "Cancer-Related Mortality and 5hmC Expression", levels = c("High", "Low"))
text(paste("P (log-rank test) = ", round(dod.med.p, digits = 3)), x = 35, y = 0.01, cex = 1.25)
```

#### Approach #2
In this approch we will plot `Kaplan-Meier curves` for overall mortality and cancer-related mortality using the upper tertile `Q3` H-score as the cutoff point for defining low/high 5hmC levels. We will compare survival curves using the `log-rank test`.

* Overall mortality
```{r kmplots2A}
# Estimating P values using the log-rank test
death.q3.lr <- survdiff(death.surv ~ q3.5hmc)
death.q3.p <- pchisq(death.q3.lr$chisq, df = 1, lower = FALSE)
# Plotting curves
survival.plot(x = survfit(death.surv ~ q3.5hmc), title = "Overall Mortality and 5hmC Expression", levels = c("High", "Low"))
text(paste("P (log-rank test) = ", round(death.q3.p, digits = 3)), x = 35, y = 0.01, cex = 1.25)
```

* Cancer-related mortality
```{r kmplots2B}
# Estimating P values using the log-rank test
dod.q3.lr <- survdiff(dod.surv ~ med.5hmc)
dod.q3.p <- pchisq(dod.med.lr$chisq, df = 1, lower = FALSE)
# Plotting curves
survival.plot(x = survfit(dod.surv ~ q3.5hmc), title = "Cancer-Related Mortality and 5hmC Expression", levels = c("High", "Low"))
text(paste("P (log-rank test) = ", round(dod.q3.p, digits = 3)), x = 35, y = 0.01, cex = 1.25)
```

***

### Logistic regression models for predicting outcome
In this section we will built logistic regression models for predicing overall mortality, cancer-related mortality, metastatic disease, and tumor progression using 5hmC expression. As before, we will categore 5hmC expression in low/high levels using as cutoff points: 1) the median; 2) the upper tertile. For each approach we will present 4 models, for each outcome: 1) unadjusted model; model adjusted by clinical features; 3) model adjusted by pathologic features; 4) model adjusted by clinical and pahtologic features.

**Clinical features** to be used for adjustment include age group, sex, and race. **Pathologic features** to be used for adjustment include high pT stage (>pT2, see below), CIS, lymphovascular invasion, and lymph node metastasis.

```{r highpt}
# Categorizing pT stages in low/high pT stage
high.pt <- ifelse(pt.cystectomy == "Tis-Ta" | pt.cystectomy == "T1" | pt.cystectomy == "T2" | pt.cystectomy == "T2a" | pt.cystectomy == "T2b", c("Low pT stage"), c("High pT stage"))
```

#### Approach #1
In this approach we will estimate `odds ratios` for overall mortality, cancer-related mortality, metastatic disease, and tumor progression using 5hmC expression categorized in low/high levels using the `median`as the cutoff point.

* Overall mortality
```{r logistic1A}
# Unadjusted model
logistic(death ~ med.5hmc, invasive)
# Model adjusted by clinical features
logistic(death ~ med.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(death ~ med.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(death ~ med.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Cancer-related mortality
```{r logistic1B}
# Unadjusted model
logistic(dod ~ med.5hmc, invasive)
# Model adjusted by clinical features
logistic(dod ~ med.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(dod ~ med.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(dod ~ med.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Metastatic disease
```{r logistic1C}
# Unadjusted model
logistic(metastasis ~ med.5hmc, invasive)
# Model adjusted by clinical features
logistic(metastasis ~ med.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(metastasis ~ med.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(metastasis ~ med.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Tumor progression
```{r logistic1D}
# Unadjusted model
logistic(death ~ med.5hmc, invasive)
# Model adjusted by clinical features
logistic(death ~ med.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(death ~ med.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(death ~ med.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

#### Approach #2
In this approach we will estimate `odds ratios` for overall mortality, cancer-related mortality, metastatic disease, and tumor progression using 5hmC expression categorized in low/high levels using the upper tertile `q3` as the cutoff point.

* Overall mortality
```{r logistic2A}
# Unadjusted model
logistic(death ~ q3.5hmc, invasive)
# Model adjusted by clinical features
logistic(death ~ q3.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(death ~ q3.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(death ~ q3.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Cancer-related mortality
```{r logistic2B}
# Unadjusted model
logistic(dod ~ q3.5hmc, invasive)
# Model adjusted by clinical features
logistic(dod ~ q3.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(dod ~ q3.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(dod ~ q3.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Metastatic disease
```{r logistic2C}
# Unadjusted model
logistic(metastasis ~ q3.5hmc, invasive)
# Model adjusted by clinical features
logistic(metastasis ~ q3.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(metastasis ~ q3.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(metastasis ~ q3.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Tumor progression
```{r logistic2D}
# Unadjusted model
logistic(progression ~ q3.5hmc, invasive)
# Model adjusted by clinical features
logistic(progression ~ q3.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
logistic(progression ~ q3.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
logistic(progression ~ q3.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

***

### Cox's proportional hazards regression models for predicting outcome
In this section we will built Cox's models for predicing overall mortality and cancer-related mortality using 5hmC expression. As before, we will categore 5hmC expression in low/high levels using as cutoff points: 1) the median; 2) the upper tertile. For each approach we will present 4 models, for each outcome: 1) unadjusted model; model adjusted by clinical features; 3) model adjusted by pathologic features; 4) model adjusted by clinical and pahtologic features. We will use `follow-up from cystectomy` for the time-to-event analysis.

**Clinical features** to be used for adjustment include age group, sex, and race. **Pathologic features** to be used for adjustment include high pT stage (>pT2, see above), CIS, lymphovascular invasion, and lymph node metastasis.

#### Approach #1
In this approach we will estimate `hazard ratios` for overall mortality, cancer-related mortality, metastatic disease, and tumor progression using 5hmC expression categorized in low/high levels using the `median`as the cutoff point.

* Overall mortality
```{r cox1A}
# Unadjusted model
hazard(death.surv ~ med.5hmc, invasive)
# Model adjusted by clinical features
hazard(death.surv ~ med.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
hazard(death.surv ~ med.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
hazard(death.surv ~ med.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Cancer-related mortality
```{r cox1B}
# Unadjusted model
hazard(dod.surv ~ med.5hmc, invasive)
# Model adjusted by clinical features
hazard(dod.surv ~ med.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
hazard(dod.surv ~ med.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
hazard(dod.surv ~ med.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```


#### Approach #2
In this approach we will estimate `hazard ratios` for overall mortality, cancer-related mortality, metastatic disease, and tumor progression using 5hmC expression categorized in low/high levels using the upper tertile `q3` as the cutoff point.

* Overall mortality
```{r cox2A}
# Unadjusted model
hazard(death.surv ~ q3.5hmc, invasive)
# Model adjusted by clinical features
hazard(death.surv ~ q3.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
hazard(death.surv ~ q3.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
hazard(death.surv ~ q3.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

* Cancer-related mortality
```{r cox2B}
# Unadjusted model
hazard(dod.surv ~ q3.5hmc, invasive)
# Model adjusted by clinical features
hazard(dod.surv ~ q3.5hmc + age.group + sex + race, invasive)
# Model adjusted by pathologic features
hazard(dod.surv ~ q3.5hmc + high.pt + cis + lvi + ln.met, invasive)
# Model adjusted by clinical and pathologic features
hazard(dod.surv ~ q3.5hmc + age.group + sex + race + high.pt + cis + lvi + ln.met, invasive)
```

***