## Data Analysis

```{r GlobalOptions}
# Setting the global options for the analysis
library(knitr)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 9)
```

```{r libraries}
# Loading the requiered libraries and sources
library(plyr)
source("StandardAnalysis3.R")
```

```{r dataset}
# Loading the full dataset
DataFull <- read.csv("../tma852-855_dataset.csv", na.strings = ".")
# Selecting the variables for the study
Variables <- c("caseid", "subtype", "grade", "X5hmc")
DataSelected <- DataFull[, Variables]
# Subsetting the data to exclude rows with missing cases
DataSpot <- subset(DataSelected, !is.na(X5hmc))
DataSpot <- subset(DataSpot, !is.na(grade))
# Renaming variable of interest
DataSpot <- rename(DataSpot, c("X5hmc" = "fhmc"))
# Renaming levels of the "subtype" variable
DataSpot$subtype <- revalue(DataSpot$subtype, c("basaloid" = "Basaloid", "papillary" = "Papillary", "usual" = "Usual", "warty" = "Warty", "warty-basaloid" = "Warty-Basaloid"))
# Converting numeric variable "grade" to factor variable and renaming levels
DataSpot$grade <- factor(DataSpot$grade)
DataSpot$grade <- revalue(DataSpot$grade, c("1" = "Grade 1", "2" = "Grade 2", "3" = "Grade 3"))
# Creating a "case-by-case" dataset
DataCase <- ddply(DataSpot, c("caseid", "subtype"), summarize, gradeCase = max(as.numeric(grade)), fhmcCase = median(fhmc))
# Converting numeric variable "gradeCase" to factor variable and renaming levels
DataCase$gradeCase <- factor(DataCase$gradeCase)
DataCase$gradeCase <- revalue(DataCase$gradeCase, c("1" = "Grade 1", "2" = "Grade 2", "3" = "Grade 3"))
# Dropping unused levels
DataSpot <- droplevels(DataSpot)
DataCase <- droplevels(DataCase)
```

### Results
A total of `r nrow(DataSpot)` TMA spots were analyzed, corresponding to `r nrow(DataCase)` cases of penile squamous cell carcinoma. The distribution by histologic subtype and grade (for all cases) is shown below.

* __Distribution by histologic subype__
```{r SubtypeDistribution}
table.prop(DataCase$subtype)
```

* __Distirbution by histologic grade__
```{r GradeDistribution}
table.prop(DataCase$gradeCase)
```

***

#### Association of 5mhC with histologic grade
```{r GradeKW}
# Spot-by-spot KW test
kruskal.test(fhmc ~ grade, data = DataSpot)
# Case-by-Case KW test
kruskal.test(fhmcCase ~ gradeCase, data = DataCase)
```

```{r GradePlots}
par(mfrow = c(1, 2))
# Spot-by-spot boxplot
with(DataSpot, boxplot(fhmc ~ grade, main = "Spot-by-Spot"))
# Case-by-case boxplot
with(DataCase, boxplot(fhmcCase ~ gradeCase, main = "Case-by-Case"))
```

***

#### Association of 5hmC with histologic subtype
```{r SubtypeKW}
kruskal.test(fhmcCase ~ subtype, data = DataCase)
```

```{r SubtypePlots, fig.height = 15, fig.width = 9}
par(mfrow = c(3, 1), cex = 1.1)
with(DataCase, boxplot(fhmcCase ~ subtype, main = "Subtype"))
with(DataCase, barplot(table(gradeCase, subtype), beside = TRUE))
with(DataCase[DataCase$subtype == "Usual",], boxplot(fhmcCase ~ gradeCase, main = "Usual SCC"))

```

